Bayesian multivariate skew-normal finite mixture model for analysis of infant development trajectories

### 2019-07-09 manuscript/supplement_old_07-09-2019.tex
sim 1

%  ALWAYS USE THE referee OPTION WITH PAPERS SUBMITTED TO BIOMETRICS!!!
%  You can see what your paper would look like typeset by removing
%  the referee option.  Because the typeset version will be in two
%  columns, however, some of your equations may be too long. DO NOT
%  use the \longequation option discussed in the user guide!!!  This option
%  is reserved ONLY for equations that are impossible to split across 
%  multiple lines; e.g., a very wide matrix.  Instead, type your equations 
%  so that they stay in one column and are split across several lines, 
%  as are almost all equations in the journal.  Use a recent version of the
%  journal as a guide. 
%  
\documentclass[useAMS,referee]{biom}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{blindtext}
\usepackage{booktabs}
\usepackage{breqn}
\usepackage{algorithm}
\usepackage{algcompatible}
\usepackage{algpseudocode}
\usepackage{float}
\usepackage{dsfont}

\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

@@

%  This will produce the submission and review information that appears
%  right after the reference section.  Of course, it will be unknown when
%  you submit your paper, so you can either leave this out or put in 
%  sample dates (these will have no effect on the fate of your paper in the
%  review process!)

\date{{\it Received October} 2007. {\it Revised February} 2008.  {\it
Accepted March} 2008.}

%  These options will count the number of pages and provide volume
%  and date information in the upper left hand corner of the top of the 
%  first page as in published papers.  The \pagerange command will only
%  work if you place the command \label{firstpage} near the beginning
%  of the document and \label{lastpage} at the end of the document, as we
%  have done in this template.

%  Again, putting a volume number and date is for your own amusement and
%  has no bearing on what actually happens to your paper!

\pagerange{\pageref{firstpage}--\pageref{lastpage}} 
\volume{64}
\pubyear{2008}
\artmonth{December}

%  The \doi command is where the DOI for your paper would be placed should it
%  be published.  Again, if you make one up and stick it here, it means 
%  nothing!

\doi{10.1111/j.1541-0420.2005.00454.x}

%  This label and the label ``lastpage'' are used by the \pagerange
%  command above to give the page range for the article.  You may have 
%  to process the document twice to get this to match up with what you 
%  expect.  When using the referee option, this will not count the pages
%  with tables and figures.

\label{firstpage}

%  put the summary for your paper here

\begin{abstract}
In studies of infant motor development, a crucial research goal is to identify latent clusters of infants that experience delayed development, as this is a known risk factor for adverse outcomes later in life. However, there are a number of statistical challenges in modeling infant development: the data are typically skewed, exhibit intermittent missingness, and are highly correlated across the repeated measurements collected during infancy. Using data from the Nurture study, a cohort of over 600 mother-infant pairs followed from pregnancy to 12 months postpartum, we develop a flexible Bayesian latent class model for the analysis infant motor development. Our model has a number of attractive features. First, we adopt the multivariate skew normal distribution with cluster-specific parameters that accommodate the inherent correlation and skewness in the data. Second, we model the cluster membership probabilities using a novel Pólya-Gamma data-augmentation scheme, thereby improving predictions of the cluster membership allocations. Lastly, we impute missing responses under missing at random assumption by drawing from appropriate conditional skew normal distributions. Bayesian inference is achieved through straightforward Gibbs sampling, and can be carried out in available software such as R.  Through simulation studies, we show that the proposed model yields improved inferences over models that ignore skewness. In addition, our imputation method yields improvements compared to conventional missing data methods, including multiple imputation and complete or available case analysis. When applied to Nurture data, we identified two distinct development clusters: one characterized by delayed “U-shaped” development and a higher percentage of male infants and another characterized by more steady development and a lower percentage of males. The clusters also differed in terms of key demographic variables, such as infant race and maternal pre-pregnancy body mass index. These findings can aid investigators in targeting interventions during this critical early-life developmental window.
\end{abstract}

%  Please place your key words in alphabetical order, separated
%  by semicolons, with the first letter of the first word capitalized,
%  and a period at the end of the list.
%

\begin{keywords}
A key word; But another key word; Still another key word; Yet another key word.
\end{keywords}

%  As usual, the \maketitle command creates the title and author/affiliations
%  display

\maketitle

\setcounter{tocdepth}{3}
\tableofcontents

@@

%  If you are using the referee option, a new page, numbered page 1, will
%  start after the summary and keywords.  The page numbers thus count the
%  number of pages of your manuscript in the preferred submission style.
%  Remember, ``Normally, regular papers exceeding 25 pages and Reader Reaction 
%  papers exceeding 12 pages in (the preferred style) will be returned to 
%  the authors without review. The page limit includes acknowledgements, 
%  references, and appendices, but not tables and figures. The page count does 
%  not include the title page and abstract. A maximum of six (6) tables or 
%  figures combined is often required.''

%  You may now place the substance of your manuscript here.  Please use
%  the \section, \subsection, etc commands as described in the user guide.
%  Please use \label and \ref commands to cross-reference sections, equations,
%  tables, figures, etc.
%
%  Please DO NOT attempt to reformat the style of equation numbering!
%  For that matter, please do not attempt to redefine anything!

\section{Introduction}
\label{s:intro}

\subsection{Infant Development Clustering}

Heterogeneity of treatment effects (HTE) (Lanza and Rhoades, 2013).

\subsection{Existing Approaches}

Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components. Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010), however in these models, focus lies primary on marginal density estimation and inference on the mixture components (i.e. clusters) is not discussed. More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed for settings in which cluster-specific inference is of primary interest (Lin \textit{et al.} 2018). However, an important feature not included in the MSTFA is the ability to explain individual-level cluster membership as a function of covariates of interest. Additionally, parameter estimation proposed by Lin et al. for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, namely the ability to incorporate prior information into a model and make posterior probability statements. Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya Gamma distribution.

\textbf{\textit{Put lit review of Bayesian PG multinomial logistic regression here}}

A ubiquitous feature of repeated measures studies is loss of data due to intermittent missingness and attrition. In the Bayesian setting, the standard approach to dealing with missing data is to perform multiple imputation, whereby $m$ imputed data sets are generated from a specified imputation model. After $m$ complete data sets are obtained, parameter estimates are combined across each data set to produce a final set of parameter estimates (Gelman \textit{et al.} 2013). This approach is not only computationally burdensome, requiring storage and analysis of an $m \times n_{rows} \times n_{cols}$ data array in addition to multiplication of total model run time by a factor of $m$, but it has been shown to produce unreliable inferences (Zhou and Reiter, 2010). We instead include an ``online" imputation imputation step in our Gibbs sampling procedure, whereby missing outcomes are updated at each iteration. This approach greatly increases the number of opportunities for exploration of the missing data parameter space.

\section{Nurture Study}
\label{s:nurt}
\subsection{Baseline Demographics and Description of Variables}

\subsection{Statistical Challenges}
\subsubsection{Skewness of Bayley score residuals}
\subsubsection{Attrition and Intermittent Missingness}

\section{Model}
\label{s:model}

\subsection{Generic Multivariate Skew Normal Mixture}

Let $\mathbf{y}_{i}$ be the $J \times 1$ observation vector for subject $i$ such that $y_{ij}$ is the observation for subject $i$ at timepoint $j$. We assume for now that $\mathbf{y}_{i}$ is fully observed. Later, we relax this assumption by allowing for missingness in the components of $\mathbf{y}_{i}$. For the analysis of the Nurture data, we propose the following MSN mixture
\begin{equation}
f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k),
\end{equation}
where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$). For now we assume that $K$ is fixed, but we discuss model selection strategies for choosing the optimal value of $K$ in Section 4. The probability that subject $i$ ($i = 1,...,n$) belongs to cluster $k$ is denoted $\pi_{ik}$. Note that $\pi_{ik}$ is indexed by $i$, as we allow the probability of class membership to vary across subjects. However, we assume a that class membership is fixed throughout the study period, since our focus is to cluster individuals based on their overall developmental patterns over the course of the study. In Section 6, we discuss extensions to allow for class membership to vary over time.

In each cluster, we assume that the $J\times 1$ outcome vector for subject $i$ follows an MSN distribution. Let $z_i \in \{1,...,K\}$ denote a latent cluster membership of subject $i$. Conditional on $z_i = k$, we model $\mathbf{y}_{i}$ as 
\begin{equation}
\mathbf{y}_{i}|z_i=k \sim MSN_J(\boldsymbol\zeta_k,\boldsymbol\alpha_k,\boldsymbol\Omega_k),
\end{equation}
where $\boldsymbol\zeta_k$ is a $J \times 1$ vector of cluster-specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster-specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster-specific scale matrix that accounts for associations between the $J$ responses. The vector $\boldsymbol\alpha_k$ has components $\alpha_{jk}$, $j = 1,...,J$, that control the skewness of outcome $j$ in cluster $k$. When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $N_J(\boldsymbol\zeta_k,\boldsymbol\Sigma_k)$, where $\boldsymbol\Sigma_J$ is the $J \times J$ covariance matrix capturing dependence among the $J$ outcomes.

\subsection{Multivariate Skew Normal Stochastic Representation}

We model the effect of covariates on longitudinal motor development outcomes through the use of a MSN regression model. We adopt a standard stochastic representation of a MSN random variable by conditioning on a subject-specific latent truncated normal random effect, denoted $t_i$ (Fr\"{u}hwirth-Schnatter and Pyne 2010). Conditional on membership in cluster $k$, the stochastic representation for the $j^{th}$ component of the multivariate skew normal distribution for subject $i$ is 
\begin{equation}
y_{ij} = \zeta_{ijk} +  \psi_{jk} t_i + \sqrt{1-\psi^2_{jk}}\epsilon_{ijk}.
\end{equation}
where $y_{ij}$ denotes the $j^{th}$ observation for subject $i$, $\zeta_{ijk}$ denotes the location parameter for the $j^{th}$ observation for subject $i$ in cluster $k$, $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a subject-specific truncated normal random effect, $\psi_{jk}$ controls the skewness of the $j^{th}$ outcome in cluster $k$, and $(\epsilon_{i1k},...,\epsilon_{iJk}) = \boldsymbol\epsilon_{ik} \sim N_J(0,\boldsymbol\Sigma_k)$ is a multivariate normal error term. Combining all observations for subject $i$ into $\mathbf{y}_i$, and conditional on $t_i$ and $z_i = k$, $\mathbf{y}_i$ follows a multivariate normal distribution with conditional mean $\mathbf{x}_i^T \boldsymbol\beta_k + t_i \boldsymbol\psi_k$ and conditional covariance $\boldsymbol\Sigma_k$. In the multivariate case, $\boldsymbol\psi_k$ is a $J \times 1$ vector containing cluster-specific skewness parameters. Marginally (after integrating over $t_i$), $\mathbf{y}_i$ is distributed $MSN_J(\boldsymbol\zeta_{ik}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where
$$\boldsymbol\zeta_{ik} = \mathbf{x}_i^T \boldsymbol\beta_k,$$
$$\boldsymbol\alpha_k = \frac{1}{\sqrt{1 - \boldsymbol\psi_k^T \boldsymbol\psi_k}} \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k,$$
$$\boldsymbol\Omega_k = \boldsymbol\Psi_k \boldsymbol\Sigma_k \boldsymbol\Psi_k + \boldsymbol\psi_k \boldsymbol\psi_k^T,$$
where $\boldsymbol\Psi_k = \text{Diag}\left ( \sqrt{1 - \psi^2_{k1}},...,\sqrt{1 - \psi^2_{kJ}} \right )$. To allow for time varying effects in our model, we structure $\mathbf{X}_i$ as a $J \times pJ$ vector of covariate values for subject $i$, and $\boldsymbol\beta_k$ as a $pJ \times 1$ vector of fixed effects coefficients for cluster $k$ (think about how to explain this better).

\subsection{Representation as Matrix Skew Normal}

We note that the multivariate skew normal density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005). As we will see in Section 3.6, this specification facilitates prior specification by admitting convenient joint prior distributions for the regression coefficients and scale matrices. In particular, we define $\mathbf{Y}_k$ as the ${n_k \times J}$  response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$). 
$$\mathbf{Y}_{k} \sim MatSN_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)$$
$$\text{vec}(\mathbf{Y}) =  (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})$$
$$\text{vec}(\mathbf{M}) = (\boldsymbol\zeta_1^T,...,\boldsymbol\zeta_{n_k}^T)$$
where $\boldsymbol\zeta_i = \mathbf{x}_i^T \boldsymbol\beta_k$ and $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})$. This implies that the $i^{th}$ row of $\mathbf{Y}_k$, a ${n_k \times J}$ matrix of responses for cluster $k$, is $\mathbf{y}_i \sim MSN_J(\boldsymbol\zeta_i, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$. Using the stochastic representation presented in Section 3.2, the conditional distribution of $\mathbf{y}_i | z_i, t_i$ is $N_J (\mathbf{x}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i,\boldsymbol\Sigma_k)$. This implies that all responses for cluster $k$ conditioned on $\mathbf{t}_{k}$ is distributed as 
$$\mathbf{Y}_k | \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),$$
where $\text{vec}(\mathbf{M}_k)_{n_kJ \times 1} = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ and $\mathbf{X}_k$ is an $n_k \times p$ design matrix.

\subsection{Multinomial Regression on Cluster Probabilities}

A primary concern of our model is with identification of latent infant development classs. We accomplish this via multinomial logit regression model on cluster membership, which utilizes P\'{o}lya-Gamma data-augmentation, as described in Section 3.6 to allow for updating of all parameters using Gibbs sampling. The multinomial logit model is as follows for $k = 1,...,K$.
$$P(z_i = k|\mathbf{w}_i) = \pi_{ik} = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{k' = 1}^K e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}}$$

where $\mathbf{w}_i$ is the vector of cluster probability covariates for subject $i$, $\boldsymbol\delta_k$ contains the multinomial regression parameters for cluster $k$, and $K$ is the number of putative classs specified \textit{a priori}. For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$. Under this model, $z_i|\mathbf{w}_i \sim Multinom(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$. During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments.

\subsection{Conditional MVN Imputation}

We allow for missingness of outcomes in the MSN mixture model by imputing missing values from their conditional multivariate normal distributions. We note that
$$\mathbf{y}_i|z_i,\mathbf{x}_i,t_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N_J(\boldsymbol\beta_{k}^T\mathbf{x}_i  + t_i \boldsymbol\psi_{k}, \boldsymbol\Sigma_{k})$$

where $z_i = k$. This allows us to appeal to standard conditional forms of the multivariate normal distribution to specify the distribution of missing observations. For subject $i$, we use $q_i$ to denote the number of the total $J$ possible repeated measurements are missing. We partition the full $J \times 1$ outcome vector $\mathbf{y}_i$ into the $(J - q_i) \times 1$ observed data vector $\mathbf{y}_i^{obs}$ and the $q_i \times 1$ missing data vector $\mathbf{y}_i^{miss}$. $Y_i = [Y^{miss}_{i_{q \times 1}} | Y^{obs}_{i_{J - q \times 1}}]^T$. We have
$$\mathbf{y}_i^{miss}|\mathbf{y}_i^{obs},\mathbf{x}_i,t_i,z_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N(\boldsymbol\mu^{miss},\boldsymbol\Sigma^{miss})$$

where $\boldsymbol\mu^{miss}$ and $\boldsymbol\Sigma^{miss}$ take standard forms detailed in the Appendix.

An attractive feature of this imputation approach is that each missing outcome is imputed ``online", i.e. once per MCMC iteration. This provides more opportunities to explore the parameter space than multiple imputation and avoids multiplicative run-time scaling in $m$, the number of imputations. We demonstrate this feature using simulations detailed in Section 4.

\subsection{Bayesian Inference}

\subsubsection{P\'olya--Gamma Data Augmentation}

Polson \textit{et al.} (2013) introduced an efficient data augmentation approach to fitting several GLMs including the multinomial logit regression model specified by Holmes and Held (2006), which specifies the full conditional distributions of the multinomial regression parameters as a function of a Bernoulli likelihood. 
$$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{k' \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}}$$

where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = \mathds{1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4. We can rewrite $\pi_{ik}$ as follows
$$\pi_{ik} = P(U_{ik} = 1) = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}}{1 + e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}} = \frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}}$$

where ${c}_{ik} = \log \sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}$ and $\eta_{ik} = \mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}$. We note that the sum $\sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}$ includes the reference category, but since we fix $\boldsymbol\delta_K = \mathbf{0}$, we have $e^{\mathbf{w}_i^T \boldsymbol\delta_K} = 1$, and hence
$$c_{ik} = \log \sum_{k' \ne k} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}} = \log \left ( 1 + \sum_{k' \notin \{k,K \}} e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}} \right )$$

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{pdflscape}
\usepackage{graphicx}
\usepackage{blindtext}
\usepackage{booktabs}
\usepackage{breqn}
\usepackage{algcompatible}
\usepackage{float}
\usepackage{dutchcal}
\usepackage{amsfonts}

@@

\subsection*{Derivations of Full Conditional Distributions}

\subsubsection*{Multivariate Skew-Normal Regression}

\subsubsection*{Multinomial Logit Regression}

\subsubsection*{Multivariate Normal Conditional Imputation}

@@

\subsection*{Tables}

\begin{algorithm}
\caption{Gibbs Sampler}
\label{alg:MCMC}
\begin{algorithmic}
\small
\linespread{0.5}
    \State $\mathbf{Define} \ n_{iter};\  n_{burn}; \ K; \ \boldsymbol\theta_{init}; \ \boldsymbol\theta_0$
    \State $n_{sim} := n_{iter} - n_{burn}$
    \State $\boldsymbol\theta := \boldsymbol\theta_{init}$
    \For {$\iota = 1,...,n_{sim}$}
        \State \textsc{I. Conditional Imputation}
          \For {$ i = 1,...,n$}
            \State $\mathbf{Draw} \ \mathbf{y}_i^{miss} \ \text{from} \ N_q(\boldsymbol\mu_i^{miss}, \boldsymbol\Sigma_i^{miss})$
            \State $\mathbf{y}_i := \mathbf{y}_i^{miss} \cup \mathbf{y}_i^{obs}$
         \EndFor
        \State \textsc{II. MSN Regression}
          \For {$ k = 1,...,K$}
            \State $n_k := \sum_{i = 1}^n {1}_{z_i = k}$
            \For {$i_k = 1,...,n_k$}
              \State $\mathbf{Draw}\  t_i \ \text{from} \ N_{[0,\infty)}(a_i,A)$
            \EndFor
            \State $\mathbf{X^*}_k := $ \texttt{cbind}$(\mathbf{X}_k,\mathbf{t}_k)$
            \State $\mathbf{Draw} \ \mathbf{B^*}_k \ \text{from} \ \text{MatNorm}(\mathbf{B}_k,\mathbf{L}_k^{-1},\boldsymbol\Sigma_k)$
            \State $\mathbf{Draw} \ \boldsymbol\Sigma_k \ \text{from} \ \text{InvWish}(\nu_k, \mathbf{V}_k)$
          \EndFor
        \State \textsc{III. Multinomial Logit}
          \For {$i = 1,...,n$}
            \For {$k = 1,...,K$}
              \State $\pi_{ik} := P(z_i = k|\mathbf{w}_i,\boldsymbol\delta_k)$
              \State $p_{ik} := P(\mathbf{y}_i|\boldsymbol\beta_k^{*T} \mathbf{x}^*_i,\boldsymbol\Sigma_k)$
            \EndFor
            \State $\mathbf{p}_{z_i} := \frac{\mathbf{p}_i \circ \boldsymbol\pi_i}{\mathbf{p}_i \cdot \boldsymbol\pi_i}$
            \State $\mathbf{Draw} \ z_i \ \text{from} \ \text{Categorical}(\mathbf{p}_{z_i})$
            \For {$k = 1,...,K-1$}
              \State $\mathbf{Draw} \ \boldsymbol\delta_k \ \text{from} \ N(\mathbf{M},\mathbf{S})$
            \EndFor
          \EndFor
        \State $\boldsymbol\theta := \{\mathbf{B^*}, \boldsymbol\Sigma, \mathbf{Z}, \boldsymbol\delta \}$
        \State $\mathbf{Store} \ \boldsymbol\theta$
	  \EndFor
\end{algorithmic}
\end{algorithm}

\begin{landscape}\begin{table}[t]

\caption{\label{tab:unnamed-chunk-5}Model results for simulated data with n = 1,000, J = 4, p = 2, K = 3, r = 2. 1,000 iterations were run with a burn in of 100. Missingness mechanism was MAR and P(miss) = 0. Model results for the multivariate skew normal (MSN) and multivariate normal (MN) mixtures are presented.}
\centering
\fontsize{6}{8}\selectfont
\begin{tabular}{lllllllllll}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Class 1} & \multicolumn{3}{c}{Class 2} & \multicolumn{3}{c}{Class 3} \\
\cmidrule(l{3pt}r{3pt}){3-5} \cmidrule(l{3pt}r{3pt}){6-8} \cmidrule(l{3pt}r{3pt}){9-11}
Component & Param. & True & MSN Est. (95\% CrI) & MN Est. (95\% CrI)  & True & MSN Est. (95\% CrI) & MN Est. (95\% CrI) & True & MSN Est. (95\% CrI) & MN Est. (95\% CrI)\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{ }}\\
\hspace{1em}MVSN & $\beta_{11}$ & 11 & 11.07 (10.74, 11.39) & 9.42 (8.91, 9.77) & -5 & -4.95 (-5.2, -4.68) & -4.11 (-4.33, -3.86) & -10 & -10.3 (-10.56, -10.01) & -8.37 (-8.86, -0.65)\\
\hspace{1em}Regression & $\beta_{21}$ & 12 & 12.02 (11.87, 12.17) & 11.98 (11.77, 12.18) & -4 & -4 (-4.1, -3.89) & -3.98 (-4.09, -3.87) & -11 & -11 (-11.19, -10.82) & -10.85 (-11.1, 0.61)\\
\hspace{1em} & $\beta_{31}$ & 13 & 13.06 (12.75, 13.36) & 11.39 (10.7, 11.78) & -3 & -2.97 (-3.25, -2.68) & -3.74 (-3.99, -3.5) & -12 & -11.9 (-12.22, -11.56) & -10.29 (-10.78, -0.14)\\
\hspace{1em} & $\beta_{41}$ & 14 & 14.06 (13.91, 14.22) & 14.02 (13.78, 14.22) & -2 & -1.96 (-2.07, -1.86) & -1.97 (-2.08, -1.84) & -13 & -13.04 (-13.25, -12.87) & -12.89 (-13.13, 0.59)\\
\hspace{1em} & $\beta_{12}$ & 2 & 2.11 (1.82, 2.35) & 0.42 (0.03, 0.83) & 5 & 5.16 (4.88, 5.47) & 5.83 (5.59, 6.07) & -2 & -1.86 (-2.21, -1.52) & -0.35 (-0.77, 0.16)\\
\hspace{1em} & $\beta_{22}$ & 2 & 2.03 (1.88, 2.17) & 2.02 (1.86, 2.22) & 5 & 4.96 (4.86, 5.06) & 4.96 (4.84, 5.07) & -2 & -1.97 (-2.18, -1.79) & -1.89 (-2.11, 0.02)\\
\hspace{1em} & $\beta_{32}$ & 2 & 2.13 (1.8, 2.43) & 0.49 (0.14, 0.86) & 5 & 5.22 (4.96, 5.49) & 4.23 (3.96, 4.5) & -2 & -1.82 (-2.14, -1.5) & -0.37 (-0.77, 0.16)\\
\hspace{1em} & $\beta_{42}$ & 2 & 2.08 (1.93, 2.23) & 2.08 (1.92, 2.28) & 5 & 4.97 (4.86, 5.08) & 4.96 (4.83, 5.08) & -2 & -1.93 (-2.13, -1.77) & -1.84 (-2.04, 0.04)\\
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{ }}\\
\hspace{1em} & $\Omega_{11}$ & 5 & 4.99 (3.84, 6.52) & 3.08 (2.69, 4.16) & 2 & 1.95 (1.52, 2.52) & 1.38 (1.2, 1.59) & 5 & 6.27 (4.84, 7.88) & 3.49 (2.89, 186.69)\\
\hspace{1em} & $\Omega_{12}$ & 4.5 & 4.55 (3.49, 5.88) & 2.76 (2.39, 3.55) & -0.5 & -0.51 (-0.83, -0.22) & 0.2 (0.05, 0.36) & 4.5 & 4.95 (3.78, 6.43) & 3.02 (2.48, 215.36)\\
\hspace{1em} & $\Omega_{13}$ & 4.25 & 4.53 (3.48, 5.85) & 1.9 (1.58, 2.46) & 1.25 & 1.08 (0.74, 1.53) & 0.38 (0.24, 0.54) & 4.25 & 4.7 (3.56, 6.04) & 1.86 (1.42, 8.74)\\
\hspace{1em} & $\Omega_{14}$ & 4.12 & 4.33 (3.31, 5.57) & 1.78 (1.46, 2.24) & -0.88 & -1.08 (-1.5, -0.75) & -0.41 (-0.58, -0.26) & 4.12 & 4.51 (3.43, 5.69) & 1.72 (1.29, 12.25)\\
\hspace{1em} & $\Omega_{22}$ & 5 & 5.09 (3.98, 6.55) & 3.43 (3.01, 4.23) & 2 & 1.99 (1.53, 2.51) & 1.43 (1.22, 1.67) & 5 & 4.87 (3.75, 6.33) & 3.71 (3.09, 253.61)\\
\hspace{1em} & $\Omega_{23}$ & 4.5 & 4.77 (3.69, 6) & 2.13 (1.78, 2.7) & -0.5 & -0.46 (-0.75, -0.18) & 0.05 (-0.1, 0.2) & 4.5 & 4.29 (3.23, 5.61) & 2.02 (1.55, 13.33)\\
\hspace{1em} & $\Omega_{24}$ & 4.25 & 4.38 (3.39, 5.68) & 1.86 (1.53, 2.39) & 1.25 & 1.45 (1.04, 1.89) & 0.64 (0.48, 0.83) & 4.25 & 4.01 (3.04, 5.19) & 1.78 (1.35, 16.8)\\
\hspace{1em} & $\Omega_{33}$ & 5 & 5.47 (4.3, 6.67) & 2.55 (2.23, 3.2) & 2 & 1.73 (1.36, 2.3) & 1.41 (1.22, 1.64) & 5 & 4.79 (3.6, 6.24) & 2.44 (1.99, 5.81)\\
\hspace{1em} & $\Omega_{34}$ & 4.5 & 4.8 (3.7, 6.06) & 1.99 (1.69, 2.68) & -0.5 & -0.68 (-1.07, -0.37) & 0.14 (-0.01, 0.3) & 4.5 & 4.25 (3.2, 5.45) & 1.94 (1.55, 5.3)\\
\hspace{1em} & $\Omega_{44}$ & 5 & 5.17 (3.96, 6.69) & 2.45 (2.14, 3.79) & 2 & 2.34 (1.8, 2.97) & 1.57 (1.37, 1.83) & 5 & 4.64 (3.54, 5.92) & 2.37 (1.93, 5.68)\\
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{ }}\\
\hspace{1em} & $\alpha_{1}$ & -0.99 & -0.81 (-2.12, 0.05) & 0 (0, 0) & 0.85 & 1.05 (0.37, 1.91) & 0 (0, 0) & 0.99 & 2.82 (1.21, 4.3) & 0 (0, 0)\\
\hspace{1em} & $\alpha_{2}$ & -0.5 & -0.22 (-1.3, 0.75) & 0 (0, 0) & -1.28 & -1.29 (-2.22, -0.5) & 0 (0, 0) & 0.5 & -0.07 (-1.14, 1.13) & 0 (0, 0)\\
\hspace{1em} & $\alpha_{3}$ & -0.5 & -0.96 (-2.14, 0.01) & 0 (0, 0) & 1.28 & 1.16 (0.47, 2.06) & 0 (0, 0) & 0.5 & 0.08 (-0.99, 1.46) & 0 (0, 0)\\
\hspace{1em} & $\alpha_{4}$ & -0.99 & -1.18 (-2.44, -0.06) & 0 (0, 0) & -0.85 & -0.91 (-1.76, -0.16) & 0 (0, 0) & 0.99 & 1.1 (0.07, 2.33) & 0 (0, 0)\\
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{ }}\\
\hspace{1em}Multinom. & $\delta_{11}$ & -0.08 & -0.07 (-0.27, 0.12) & -0.54 (-0.77, -0.32) & -0.08 & -0.07 (-0.27, 0.12) & -0.54 (-0.77, -0.32) & -0.08 & -0.07 (-0.27, 0.12) & -0.54 (-0.77, -0.32)\\
\hspace{1em} & $\delta_{12}$ & 0.51 & 0.25 (-0.04, 0.53) & -0.26 (-0.6, 0.05) & 0.51 & 0.25 (-0.04, 0.53) & -0.26 (-0.6, 0.05) & 0.51 & 0.25 (-0.04, 0.53) & -0.26 (-0.6, 0.05)\\
\hspace{1em} & $\delta_{21}$ & -0.97 & -0.71 (-0.95, -0.48) & -0.07 (-0.28, 0.14) & -0.97 & -0.71 (-0.95, -0.48) & -0.07 (-0.28, 0.14) & -0.97 & -0.71 (-0.95, -0.48) & -0.07 (-0.28, 0.14)\\
\hspace{1em} & $\delta_{22}$ & 0.84 & 0.39 (0.09, 0.71) & 0.24 (-0.04, 0.5) & 0.84 & 0.39 (0.09, 0.71) & 0.24 (-0.04, 0.5) & 0.84 & 0.39 (0.09, 0.71) & 0.24 (-0.04, 0.5)\\
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{ }}\\
\hspace{1em}Clustering & $\pi_l$ & 0.38 & 0.38 (0.38, 0.38) & 0.38 (0.13, 0.41) & 0.4 & 0.39 (0.39, 0.4) & 0.39 (0.36, 0.43) & 0.23 & 0.23 (0.22, 0.23) & 0.23 (0.2, 0.44)\\
\bottomrule
\end{tabular}
\end{table}
\end{landscape}

\newpage
\subsection*{Figures}

\begin{figure}[h]
	\label{fig:skew_resids}
	\caption{The distribution of residuals in repeated measures regression model of Bayley composite scores adjusted for race and sex.}
	\centering
	\includegraphics[width = 1\textwidth]{bayley_resids_plot}
\end{figure}

\begin{figure}[h]
	\label{fig:sim1_dens}
	\caption{Marginal densities of simulated outcomes in each cluster at each measurement occasion.}
	\centering
	\includegraphics[width = 1\textwidth]{y_densities.pdf}
\end{figure}

\begin{figure}[h]
	\label{fig:sim1_means}
	\caption{Mean plots of simulated outcomes in each cluster at each measurement occasion.}
	\centering
	\includegraphics[width = 1\textwidth]{y_means.pdf}
\end{figure}

\newpage
\subsection*{Glossary of Notation}
